---
title: "R Notebook"
output: html_notebook
---


```{r}
library(rstan)
library(bayesplot)
```

#Before I start model- 
I will need to edit the dataset such that I have columns for win/lose streak at every week. 

```{r}
setwd("~/Honours/Model/Model Fitting/Player Rest/Data")
#library(readr)
nba <- read.csv(file = "daysofrest.csv", header = TRUE)
# 
# nba$score_diff <- nba$H.Score - nba$A.Score
# 
# nba$nteams <- length(unique(nba$Home.Team))
# nba$ngames <- length(nba$score_diff)
# nba$nweeks <- floor(2*nba$ngames/nba$nteams) 
```

#STAN code
this is the stan code for incorporating the days of rest the home and away team has had prior to their game. 

I have used a classification for 1,2,3, 4 of more days of rest and estimated coefficients for each. 

I have used a classification of groups of win or lose streak for the home and away team. Since wining or losing 1 and 2 games wil have a small and approx same effect we will group them into one ("homeawayW2") as well as win/lose 3 & 4 and win/lose 5 or more into one group. 

**I have not combined the unique home effect with the team rest here. 
```{stan output.var="nba_model"}
data{
int<lower=1> nteams; 
int<lower=1> ngames;
int<lower=1> nweeks;
int<lower=0, upper=1> resthome2[ngames]; //2 days of rest of home team
int<lower=0, upper=1> resthome3[ngames]; //3 days of rest for home team
int<lower=0, upper=1> resthome4[ngames]; // 4 or more days of rest for home team
int<lower=0, upper=1> restaway2[ngames]; // 2 days rest away team
int<lower=0, upper=1> restaway3[ngames]; // 3 day of rest for away team
int<lower=0, upper=1> restaway4[ngames]; // 4 or more days of rest for away team

//int<lower=0, upper=1> homewin1[ngames]; // 1 game win streak home
int<lower=0, upper=1> homewin2[ngames]; // 2 or 1 game win streak home
//int<lower=0, upper=1> homewin3[ngames]; // 3 game win streak home
int<lower=0, upper=1> homewin4[ngames]; // 4 or 3 game win streak home
int<lower=0, upper=1> homewin5[ngames]; // 5 or more game win streak home
//int<lower=0, upper=1> homelose1[ngames]; // 1 game lose streak home
int<lower=0, upper=1> homelose2[ngames]; // 2 or 1game lose streak home
// int<lower=0, upper=1> homelose3[ngames]; // 3 game lose streak home
int<lower=0, upper=1> homelose4[ngames]; // 4 or 3 game lose streak home
int<lower=0, upper=1> homelose5[ngames]; // 5 or more game lose streak home

//int<lower=0, upper=1> awaywin1[ngames]; // 1 game win streak away
int<lower=0, upper=1> awaywin2[ngames]; // 2 game win streak away
//int<lower=0, upper=1> awaywin3[ngames]; // 3 game win streak away
int<lower=0, upper=1> awaywin4[ngames]; // 4 or 3game win streak away
int<lower=0, upper=1> awaywin5[ngames]; // 5 or more game win streak away
//int<lower=0, upper=1> awaylose1[ngames]; // 1 game lose streak away
int<lower=0, upper=1> awaylose2[ngames]; // 2 game lose streak away
//int<lower=0, upper=1> awaylose3[ngames]; // 3 game lose streak away
int<lower=0, upper=1> awaylose4[ngames]; // 4 game lose streak away
int<lower=0, upper=1> awaylose5[ngames]; // 5 or more game lose streak away

int<lower=1> home_week[ngames]; // week number for the home team
int<lower=1> away_week[ngames]; // week number for the away team
int<lower=1, upper=nteams> home_team[ngames]; // home team ID (1, ..., 20)
int<lower=1, upper=nteams> away_team[ngames]; // away team ID (1, ..., 20)
vector[ngames] score_diff; // home_goals - away_goals

}
parameters {
real b_home; // the effect of hosting the game in mean of score_diff dist.
//real b_prev; // regression coefficient of prev_perf


//real beta1; //regression coefficient of days of rest home
real beta2h; //regression coefficient of days of rest aways
real beta3h;
real beta4h;
real beta2a;
real beta3a;
real beta4a;

//real alphaW1H; // regression coefficient for win/loss streak away
real alphaW2H;
//real alphaW3H;
real alphaW4H;
real alphaW5H;
//real alphaL1H;
real alphaL2H;
//real alphaL3H;
real alphaL4H;
real alphaL5H;
//real alphaW1A; // regression coefficient for win/loss streak for away team
real alphaW2A;
//real alphaW3A;
real alphaW4A;
real alphaW5A;
//real alphaL1A;
real alphaL2A;
//real alphaL3A;
real alphaL4A;
real alphaL5A;


real<lower=0> sigma_a0; // teams ability variation
real<lower=0> tau_a; // hyper-param for game-to-game variation
real<lower=1> nu; // t-dist degree of freedom
real<lower=0> sigma_y; // score_diff variation
row_vector<lower=0>[nteams] sigma_a_raw; // game-to-game variation
matrix[nweeks,nteams] eta_a; // random component
}

transformed parameters {
matrix[nweeks, nteams] a; // team abilities
row_vector<lower=0>[nteams] sigma_a; // game-to-game variation
vector[ngames] a_diff;
vector[ngames] mu;
a[1] =  sigma_a0 * eta_a[1]; // initial abilities (at week 1)
sigma_a = tau_a * sigma_a_raw; //performance evolves from week to week
  for (w in 2:nweeks) {
    a[w] = a[w-1] + sigma_a .* eta_a[w]; // evolution of abilities
  }
  for (g in 1:ngames) {
    a_diff[g] = a[home_week[g],home_team[g]] - a[away_week[g],away_team[g]];
    // additional reg parameters e.g. rest of home, rest of away team
    mu[g] = a_diff[g] + b_home  + beta2h*resthome2[g] + beta3h*resthome3[g] + beta4h*resthome4[g] + beta2a*restaway2[g] + beta3a*restaway3[g] + beta4a*restaway4[g]
     + alphaW2H*homewin2[g] + alphaW4H*homewin4[g] + alphaW5H*homewin5[g] + alphaL2H*homelose2[g] 
     + alphaL4H*homelose4[g] + alphaL5H*homelose5[g] + 
    alphaW2A*awaywin2[g]+  + alphaW4A*awaywin4[g] + alphaW5A*awaywin5[g] + alphaL2A*awaylose2[g] +  
     alphaL4A*awaylose4[g] + alphaL5A*awaylose5[g] ;
    //(come in via the data section)
  }
}

model {

// Priors
nu ~ gamma(2,0.1);
//b_prev ~ normal(0,1);

beta2h ~ normal(0,10); //prior on beta1
beta3h ~ normal(0,10); //prior on beta2
beta4h ~ normal(0,10);
beta2a ~ normal(0,10);
beta3a ~ normal(0,10);
beta4a ~ normal(0,10);

sigma_a0 ~ normal(0,1);
sigma_y ~ cauchy(0,5); 
b_home ~ normal(0,100); //blangiardo uses N(0,0.0001)
sigma_a_raw ~ normal(0,1);
tau_a ~ cauchy(0,1);
to_vector(eta_a) ~ normal(0,1);
// Likelihood
score_diff ~ student_t(nu, mu, sigma_y);
}

generated quantities {
vector[ngames] score_diff_rep;
for (g in 1:ngames)
score_diff_rep[g] = student_t_rng(nu, a[home_week[g], home_team[g]] -a[away_week[g],away_team[g]]+b_home, sigma_y);
}

```

##Fitting the model

```{r}
nba <- read.csv(file = "data_withstreaks.csv", header = TRUE)
#THE DATASET ATTACHED HAS TREATED NaN AS 0S
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

nsamples <- 2000
dat <- NULL
dat$nteams <- 30
dat$ngames <- 1230
dat$nweeks <- 82
dat$home_team <- nba$H.Team.ID
dat$away_team <- nba$Away.Team.ID
dat$home_week <- nba$home_week
dat$away_week <- nba$away_week
dat$score_diff <- nba$H.Score - nba$A.Score

dat$resthome2 <- as.numeric(nba$DaysofRest.Home.Team == 2)
dat$resthome3 <- as.numeric(nba$DaysofRest.Home.Team == 3)
dat$resthome4 <- as.numeric(nba$DaysofRest.Home.Team == 0 | nba$DaysofRest.Home.Team > 3)
dat$restaway2 <- as.numeric(nba$DaysofRest.Away.Team == 2)
dat$restaway3 <- as.numeric(nba$DaysofRest.Away.Team == 3)
dat$restaway4 <- as.numeric(nba$DaysofRest.Away.Team == 0 | nba$DaysofRest.Home.Team > 3)

dat$homewin2 <- as.numeric(nba$H_Streak == "W1" | nba$H_Streak =="W2")
dat$homewin4 <- as.numeric(nba$H_Streak == "W3"| nba$H_Streak == "W4" ) 
dat$homewin5 <- as.numeric(nba$H_Streak != "W1" | nba$H_Streak != "W2" | nba$H_Streak != "W3" | nba$H_Streak != "W4")
dat$homelose2 <- as.numeric(nba$H_Streak == "L1" | nba$H_Streak == "L2")
dat$homelose4 <- as.numeric(nba$H_Streak == "L3" | nba$H_Streak == "L4")
dat$homelose5 <- as.numeric(nba$H_Streak != "L1" | nba$H_Streak != "L2" | nba$H_Streak != "L3" | nba$H_Streak != "L4")

dat$awaywin2 <- as.numeric(nba$A_Streak == "W1" | nba$A_Streak =="W2")
dat$awaywin4 <- as.numeric(nba$A_Streak == "W3"| nba$A_Streak == "W4" ) 
dat$awaywin5 <- as.numeric(nba$A_Streak != "W1" | nba$A_Streak != "W2" | nba$A_Streak != "W3" | nba$A_Streak != "W4")
dat$awaylose2 <- as.numeric(nba$A_Streak == "L1" | nba$A_Streak == "L2")
dat$awaylose4 <- as.numeric(nba$A_Streak == "L3" | nba$A_Streak == "L4")
dat$awaylose5 <- as.numeric(nba$A_Streak != "L1" | nba$A_Streak != "L2" | nba$A_Streak != "L3" | nba$A_Streak != "L4")

fit <- sampling(nba_model, chains = 4, iter = (nsamples/2), data = dat)

vectofOBS <- dat$score_diff  
```

```{r}
library(shinystan)
launch_shinystan(fit)
```

#Parameter Estimates
```{r}
sum_fit <- summary(fit)
sf <- sum_fit$summary
sf
```

#Model Checking

Check weather the actual score differences are consistent with the distribution of the replicated date. 
All 1230 games are scorted according to their score differences (shown in black dots). 
For each game, the median of the replicate score differences is shown in red, the 95% uncertainty interval is shown in light yellow, and the 50% uncertainty interval is shown in dark yellow.
```{r}
install.packages("matrixStats")
library(matrixStats)
#fit <- readRDS("FITS/fit_38.rds")
sims <- extract(fit)
scd <- dat$score_diff
scd_sims <- sims$score_diff_rep
scd_hat <- colMedians(scd_sims)
alpha <- 0.95
scd_ub <- colQuantiles(scd_sims, probs = 1-(1-alpha)/2)
scd_lb <- colQuantiles(scd_sims, probs = (1-alpha)/2)
ci95 <- sum(scd < scd_ub & scd_lb<scd)/380
alpha <- 0.5
scd_ub <- colQuantiles(scd_sims, probs = 1-(1-alpha)/2)
scd_lb <- colQuantiles(scd_sims, probs = (1-alpha)/2)
ci50 <- sum(scd < scd_ub & scd_lb<scd)/380
```

```{r}
library(ggplot2)
#library(matrixStats)
#fit <- readRDS("FITS/fit_38.rds")
sims <- extract(fit)
scd <- dat$score_diff
scd_sims <- sims$score_diff_rep
scd_hat <- colMedians(scd_sims)
scd_se <- sqrt(colVars(scd_sims))
alpha <- 0.95;
scd_ub <- colQuantiles(scd_sims, probs = 1-(1-alpha)/2)
scd_lb <- colQuantiles(scd_sims, probs = (1-alpha)/2)
alpha <- 0.5;
scd_ub2 <- colQuantiles(scd_sims, probs = 1-(1-alpha)/2)
scd_lb2 <- colQuantiles(scd_sims, probs = (1-alpha)/2)

sort_scd <- scd[order(scd)]
sort_scd_hat <- scd_hat[order(scd)]
sort_scd_se <- scd_se[order(scd)]
sort_scd_ub <- scd_ub[order(scd)]
sort_scd_lb <- scd_lb[order(scd)]
sort_scd_ub2 <- scd_ub2[order(scd)]
sort_scd_lb2 <- scd_lb2[order(scd)]
df <- data.frame(list(scd = sort_scd, scd_hat = sort_scd_hat, scd_se = sort_scd_se, 
                      scd_ub = sort_scd_ub, scd_lb = sort_scd_lb, 
                      scd_ub2 = sort_scd_ub2, scd_lb2 = sort_scd_lb2))

ggplot(df, aes(x = c(1:1230))) +
  geom_ribbon(aes(ymin = scd_lb,
                  ymax = scd_ub),
              fill="lightyellow") + 
  geom_ribbon(aes(ymin = scd_lb2,
                  ymax = scd_ub2),
              fill="khaki3") + 
  geom_line(aes(y=scd_hat),colour="darkred") + 
  geom_point(aes(y=scd), size = 0.3) +
  scale_x_continuous(name="match") +
  scale_y_continuous(name="score difference", minor_breaks = seq(-6, 6, 1), 
                     sec.axis = dup_axis()) +
  ggtitle("Estimated score differences using Player Rest Feature")

```
In the figure above, all 1230 matches are sorted according to their score difference (shown in black
dots). For each match, the median of the replicated score differences is shown in red, the
95% uncertainty interval is shown in light yellow, and the 50% uncertainty interval is shown
in dark yellow. 